---
title: "Taxonomic processing"
author: "Anna Krystalli"
date: "22/06/2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

## functions

```{r}
library(dplyr)
get_taxo_base <- function(x){tail(na.omit(unlist(x)), 1)}

get_valid_taxon <- function(base){
    qres <- try(worrms::wm_records_names(base), silent = T)
    if(class(qres) == "try-error") {
        warning("No valid_name returned for base: ", base)
        return(data.frame(valid_name = NA, 
                          aphiaID = NA))
    }else{
        return(data.frame(valid_name = qres[[1]]$valid_name[1], 
                          aphiaID = qres[[1]]$valid_AphiaID[1]))
    }
}

get_valid_id <- function(base){
    qres <- try(worrms::wm_name2id(base), silent = T)
    if(class(qres) == "try-error") {
        warning("No aphiaID returned for base: ", base)
        return(data.frame(aphiaID = NA))
    }else{
        return(data.frame(aphiaID = qres))
    }
}


worrms_taxonomy <- function(taxonomy, 
                            method = c("record", "id")){
    method <- match.arg(method)
    switch(method,
           "record" = {
               taxonomy %>% 
                   do(get_valid_taxon(.$base)) %>% 
                   bind_cols(taxonomy, .)},
           "id" = {
               taxonomy %>% 
                   do(get_valid_id(.$base)) %>% 
                   bind_cols(taxonomy, .)}
           )
}
```


## read data

Read in data and extract `base` taxon to search WoRMS for `aphiaID`.
```{r}
taxonomy <- readr::read_tsv(here::here("data", "metadata", "taxonomy.tsv")) %>%
    rowwise %>% 
    do({
        result = as_data_frame(.)
        result$base = get_taxo_base(.) %>% unlist
        result}) 
taxonomy

```

## match prey taxonomic data to WoRMS

Search WoRMS for `base` and append `valid_name` and `aphiaID` columns to `taxonomy`.

```{r, cache=TRUE}
taxonomy <- taxonomy %>% worrms_taxonomy()
```

```{r}

taxonomy %>% select(base:aphiaID)
```

### invalid names

```{r}
na_worrms <- taxonomy %>% filter(is.na(valid_name)) 
na_worrms %>% select(base:aphiaID)
```

```{r}
na_worrms %>% pull(base)
```

### odd behaviour

#### `"record"` method returns NA 
```{r}
na_worrms %>% 
    filter(base == "Gammaracanthus caspius") %>%
    pull(aphiaID)
```

where `worrms::wm_name2id()` with with the same name returns a valid `aphiaID`


```{r}
worrms::wm_name2id("Gammaracanthus caspius")
```

## test method `"id"`

```{r, cache=TRUE}
taxonomy_id <- taxonomy %>% worrms_taxonomy(method = "id")
```

- `aphiaID` is with `method = "record"`
- `aphiaID1` is with `method = "id"`

#### `"id"` method returns `-999` values

For species for which method `"record"` returns a valid record

```{r}
taxonomy_id %>% select(base:aphiaID1) %>%
    filter(aphiaID1 < 0)
```

But no NAs

```{r}
taxonomy_id %>% select(base:aphiaID1) %>%
    filter(is.na(aphiaID1))
```

and has filled a couple NAs not returned through `"method"` query.

```{r}
taxonomy_id %>% select(base:aphiaID1) %>%
    filter(is.na(aphiaID))
```

```{r}
taxonomy_id %>% select(base:aphiaID1) %>%
    filter(aphiaID1 < 0)
```

### try composite function of both methods

```{r function2}
get_valid_taxon <- function(base){
    qres <- try(worrms::wm_records_names(base), silent = T)
    if(class(qres) == "try-error") {
        aphiaID <- try(worrms::wm_name2id(base), silent = T)
        if(class(aphiaID) == "try-error") {
            warning("No valid aphiaID returned for base: ", base)
            return(data.frame(valid_name = NA, aphiaID = NA))}
        if(aphiaID < 0) {
            warning("No valid aphiaID returned for base: ", base)
            return(data.frame(valid_name = NA, aphiaID = NA))}
        qres <- worrms::wm_record(aphiaID)
        if(class(qres) == "try-error") {
            return(data.frame(valid_name = NA, aphiaID = NA))
        }
        return(data.frame(valid_name = qres$valid_name, 
                          aphiaID = qres$valid_AphiaID))
    }else{
        aphiaID <- qres[[1]]$valid_AphiaID[1]
        return(data.frame(valid_name = qres[[1]]$valid_name[1], 
                          aphiaID = ifelse(aphiaID < 0, NA, aphiaID)))
    }
}

```

```{r, cache=TRUE}
na_worrms <- readr::read_tsv(here::here("data-raw", "taxonomy.tsv")) %>%
    rowwise %>% 
    do({
        result = as_data_frame(.)
        result$base = get_taxo_base(.) %>% unlist
        result}) %>% 
    worrms_taxonomy() %>% 
    select(base:aphiaID) %>% 
    filter(is.na(aphiaID))
```

```{r}
na_worrms %>% select(base:aphiaID)
```


```{r, eval=F}
na_worrms %>% 
    select(base) %>% 
    mutate(verified_base = NA) %>%
    readr::write_csv(here::here("data-raw", "validation", 
                                paste0(format(Sys.time(), 
                                              "%Y-%m-%d_"),
                                       "worrms_nomatches.csv")))
```
