---
title: "Bibliography processing"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Load data

```{r}
library(dplyr)
load(here::here("data", "seabirddiet.rda"))
```

## Manual corrections

Required to deal with:

- multiple references in a single row
- drag down errors
- other typos

### create manual edit table

Create editable table of references to record any changes to be made to raw data values. Try to identify multiple references using a couple of regex expressions.

```{r}
# setup manual correction table
seabirddiet %>% 
    select(reference) %>% 
    distinct() %>% 
    mutate(rename = reference, 
           multi = stringr::str_detect(reference, "([0-9]+-[0-9]+,) | ([0-9]+-[0-9]+ &+ [A-z])")) %>%
    # render data table
    DT::datatable(class = 'cell-border stripe', rownames = F, 
                  editable = T,
                  extensions = 'Buttons', 
                  options = list(
                      dom = 'Bfrtip',
                      buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
                  )) %>%
    DT::formatStyle(
        'multi',
        backgroundColor = DT::styleEqual(
            c(0, 1), c('white', 'lightpink')
        )
    )

```

<br>

### make manual corrections

In particular:

- **delimit multiple references in single cell with `;`**. 
- Correct excel drag down error in references eg: `Daunt & Wanless IMLOT 1973-2015:2127`.

### save table

Save as **CSV** using table widget at path **`r here::here("data-raw", "validation", "reference_manual_clean.csv")`**



```{r}
ref_manual <- readr::read_csv(here::here("data-raw", 
                                         "validation", 
                                         "reference_manual_clean.csv")) %>%  
    mutate(rename = stringr::str_replace(rename, "–", "-")) %>%
    mutate(rename = stringr::str_replace(rename, "\\.$", ""))

```
### check manual corrections
```{r}
ref_manual %>%
    filter(reference != rename) %>%
    DT::datatable()
```

### separate references and gather into long format

Easiest way to separate is to use `tidyr::separate` and `tidyr::gather()`. Also I'll use a function to mark identifiers to link back to the original data.

```{r}

get_ref_code <- function(x, type = c("ref_tab", "raw"), width = 3){
    type <- match.arg(type)
    prefix <- switch(type,
                     "ref_tab" = "REF",
                     "raw" = "RAW")
    glue::glue("{prefix}{stringr::str_pad(x, {width}, pad = 0)}")
}

```



#### determine number of columns to spread to.

```{r}
n_ref <- stringr::str_count(ref_manual$rename, ";") %>% max() + 1
```


#### separate and gather

```{r}
ref_sep <- ref_manual %>% 
    mutate(raw_id = get_ref_code(1:nrow(.), type = "raw")) %>%
    tidyr::separate(rename, into = as.character(1:n_ref), sep = ";") %>%
    tidyr::gather(key = ref_n, value = ref_valid, as.character(1:4)) %>%
    filter(!is.na(ref_valid)) %>%
    mutate(ref_valid = stringr::str_trim(ref_valid)) 

ref_sep %>%    
    DT::datatable()
```




```{r}
ref <- ref_sep %>% distinct(ref_valid) %>%
    mutate(ref_id = get_ref_code(1:nrow(.), type = "ref")) %>%
    select(ref_id, ref_valid) %>%
    distinct()  

ref
```

### check for distinctness of references

Return to editable manual edit table if need be.

```{r}
ref_sep %>% 
    select(reference, ref_valid) %>% 
    distinct(ref_valid) %>% 
    arrange(ref_valid) %>%
    DT::datatable(editable = T)
```

***

## Process references



## get vector of refs


```{r}
refs <- ref_sep$ref_valid %>% unique()
```


```{r}
refs %>% head(15)
```

## get crossref record

Because what we have of references is the citation information, we use a bibliographic field query through argument `flq`.

```{r, cache=TRUE}
dois <- refs %>% purrr::map_df(
    ~rcrossref::cr_works(flq = c(query.bibliographic = .x),
                         limit = 5, 
                         .progress="text",
                         sort = "score")$data)
```

```{r}

ref_tbl <- bind_cols(tibble(reference = rep(refs, each = 5)), dois) %>% 
    left_join(distinct(seabirddiet, reference, source),
              by = "reference") %>% 
    select(reference, 
           author, title, everything())
```


```{r}
DT::datatable(ref_tbl, class = 'cell-border stripe', rownames = F, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
))
```

***

## Query using mutlple fields
### split refs into components

#### authors
```{r}
auth <- stringr::str_extract(refs, "^[^[:digit:]]+") %>% stringr::str_trim()
auth %>% head
```

#### date
```{r}
date <- stringr::str_extract(refs, "[:digit:]{4}+") %>% stringr::str_trim()
date

```

#### container
```{r}
container <- stringr::str_extract(refs, "(?<=[:digit:]) [^[:digit:]]+") %>% 
    stringr::str_trim() %>% stringr::str_replace(", No.", "") 

container %>% head()

```

### Functions to split the references into fields

```{r}
extr_auth <- function(x){
    stringr::str_extract(x, "^[^[:digit:]]+") %>% 
        stringr::str_trim()}

extr_year <- function(x){
    stringr::str_extract(x, "[:digit:]{4}+") %>% 
        stringr::str_trim()}

extr_container <- function(x){
    stringr::str_extract(x, "(?<=[:digit:]) [^[:digit:]]+") %>% 
        stringr::str_trim() %>% stringr::str_replace(", No.", "") 
}

extr_pages <- function(x){
    stringr::str_extract(x, "([:digit:]+-[:digit:]+)") %>% 
        stringr::str_trim() 
}


get_crfields_table <- function(refs){
    tibble(reference = refs) %>%
        mutate(author = extr_auth(reference),
               year = extr_year(reference),
               container = extr_container(reference),
               pages = extr_pages(reference))
}
```

```{r}
crfields <- get_crfields_table(refs)

DT::datatable(crfields)
```




```{r}

crq_query_fail <- function(x){
    tibble::tibble(reference = x$reference, 
                   author = NA,
                   author_f = NA,
                   year = NA,
                   container = NA,
                   volume = NA,
                   issue = NA,
                   pages = NA,
                   doi = NA,
                   title = NA
    )
}

crq_query_extr <- function(x, q){
    tibble::tibble(
        reference = x$reference, 
        author = crq_auth_conc(q$author),
        author_f = crq_auth_first(q$author),
        year = lubridate::as_date(q$created) %>% 
            lubridate::year(),
        container = q$container.title,
        volume = if(any(names(q) %in% "volume")){q$volume}else{NA},
        issue = if(any(names(q) %in% "issue")){q$issue}else{NA},
        pages = q$page,
        doi = q$doi,
        title = q$title,
        r_author = x$author,
        r_author_f = stringr::word(x$author, 1, 1),
        r_year = lubridate::as_date(x$year, format = "%Y", 
                                    tz = "GMT") %>% 
            lubridate::year(),
        r_container = x$container,
        r_pages = x$pages)
}   

crq_auth_conc <- function(author){
    author %>% purrr::map_chr(~.x %>% pull(family) %>% paste0( collapse = "; "))
}

crq_auth_first <- function(author){
    author %>% purrr::map_chr(~.x %>% filter(sequence == "first") %>% 
                                  slice(1) %>% pull(family))
}

crq_validate_ref <- function(q.out){
    q.out %>% filter(r_author_f == author_f, 
                     r_year == year,
                     #r_pages == pages
                     )
}


query_works <- function(x, limit = 5){
    q <- rcrossref::cr_works(
        flq = c(`query.container-title` = x$container[1],
                query.author = x$author[1],
                query.bibliographic =  lubridate::as_date(as.numeric(x$year[1]))),
        limit = limit)$data
    
    if(nrow(q) == 0){
        return(crq_query_fail(x))
    }
    
    q.extr <-  crq_query_extr(x, q) %>% 
        crq_validate_ref()
    
    if(nrow(q.extr) == 0){
        return(crq_query_fail(x))
    }else{
        q.extr %>% select(-starts_with("r_")) 
    }
}

```


```{r}
crfields[c(1:2, 23),]
```



```{r}
rbind(query_works(crfields[1,]),
      query_works(crfields[2,]),
      query_works(crfields[23,]))
```
```{r, message=F}
out <- crfields[,] %>% 
    rowwise %>%
    do({query_works(., limit = 10)})
    
```

Clearly not worth any of this effort for the returns

```{r}
out
```


<div class="alert alert-danger">

## ABANDON `rcrossref` matching 

</div>

### `rcrossref earlier experimentation

Try without `et al`

```{r}
rcrossref::cr_works(
    flq = c(`query.container-title` = crfields$container[2],
            query.author = "Wilson",
            query.bibliographic = lubridate::as_date(2004)
    ))
```
```{r}
rcrossref::cr_works(flq = c(query.title = "Self-feeding and chick provisioning diet differ in the Common Guillemot Uria aalge",
                            query.author = "Wilson"))


rcrossref::cr_works(flq = c(`query.container-title` = crfields$container[2],
                            query.title = "Self-feeding and chick ",
                            query.author = "Wilson"))
```



```{r, cache=TRUE}
dois <- refs %>% purrr::map_df(
    ~rcrossref::cr_works(flq = c(query.bibliographic = .x),
                         limit = 5, 
                         .progress="text",
                         sort = "score")$data)
```

***

# Issues

### `Exp.` not matched to `Experimental` in `rcrossref`

References with `Exp.` in the citation does not return correct reference
```{r}
rcrossref::cr_works(flq = c(query.bibliographic = 
                                "Blake BF 1984 Journal. Exp. Mar. Biol. Ecol. 76, 89-103"), 
                    limit = 1, 
                    sort = "score")$data
```
Changing `Exp.` to `Experimental.` in the citation returns the correct reference

```{r}
rcrossref::cr_works(flq = c(query.bibliographic = 
                                "Blake BF 1984 Journal. Experimental. Mar. Biol. Ecol. 76, 89-103"), 
                    limit = 1, 
                    sort = "score")$data
```



## 

```{r}
rcrossref::cr_works(flq = c(query.bibliographic = "Diet and fish stock availability"), 
                    limit = 1, 
                    sort = "score")$data

```


